### 2020-10-08 backup/unify/main.tex

\documentclass[12pt]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{amssymb}
\usepackage{bbm}

\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{url}

%%% For figures
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%% For tables
\usepackage[table]{xcolor}

%specifying table and figure packages
\usepackage{caption}

\usepackage[T1]{fontenc}
\usepackage[all]{xy}
\usepackage[inline]{enumitem}

\usepackage{tikz-cd}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

\crefname{diagram}{diag.}{diags.}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2(#1)#3}

\title{Weight sharing via comonads}
\author{
    Pietro Vertechi \and Mattia G. Bergomi
}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}

One of the key reasons for the success of neural networks is the notion of {\em weight sharing}. In its simplest form, one can consider, among maps $[M, X] \rightarrow [M, X]$, those that originate from maps $[M, X] \rightarrow X$. Indeed, given
\begin{equation*}
    {\cal F} \colon [M, X] \rightarrow X,
\end{equation*}
one can consider the map
\begin{align*}
    \widehat{\cal F} \colon [M, X]       & \rightarrow [M, X]              \\
    \left(\widehat{\cal F}\phi\right)(m) & = {\cal F}(a \mapsto \phi(ma)).
\end{align*}

This simple notion of weight sharing does not seem sufficient to cover complex neural networks, made up of several nodes, each with their own symmetries, which interact with each other in non-trivial ways.

We aim to show that the a similar strategy, based on a different comonad, is sufficient to recover, among other examples, classical recurrent and convolutional neural networks, incorporating at the same time equivariance and locality.

\section{Architecture}

\begin{definition}\label{def:cartesian_category}
    We say that a category is {\em Cartesian} if it has finite products. We say that a functor between Cartesian categories is {\em Cartesian} if it preserves finite products.
\end{definition}

\begin{definition}\label{def:nilpotent_functor}
    A functor $F$ is {\em nilpotent} if there exists $n \in \N$ such that the composed functor $F^n$ maps all objects to the terminal object.
\end{definition}

We begin by giving an abstract definition of {\em nilpotent architecture}.

\begin{definition}\label{def:nilpotent_architecture}
    Let $\Cat$ be a Cartesian category. A {\em nilpotent architecture} is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$.
\end{definition}

This is sufficient to give notions of composability, weight sharing, and locality.
Graph-like structures are a source of examples of nilpotent architecture.

\begin{proposition}\label{prop:graph_architecture}
    \pietro{Il multigrafo è più interessante, ma tanto è solo un esempio.}
    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a Cartesian category. The functor
    \begin{align*}
        F \colon \DCat^V & \rightarrow \DCat^V                  \\
        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
    \end{align*}
    is a nilpotent architecture.
\end{proposition}

\begin{proof}
    It is straightforward to verify that $F$ is Cartesian.
    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
\end{proof}

\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.

\begin{definition}\label{def:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a nilpotent architecture.
\end{proposition}

\begin{proof}
    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$, hence it preserves limits and is therefore Cartesian. By~\cref{eq:nilpotent_left_adjoint}, $(t_*s^*)^n$ maps all objects to the initial object, so its adjoint $(s_!t^*)^n$ maps all objects to the terminal object.
\end{proof}

\subsection{The category of coalgebras}

Let $\Cat$ be a Cartesian closed category, and let $F\colon \Cat \rightarrow \Cat$ be a nilpotent architecture. We can consider $\FCoalg$, the category of coalgebras over $F$ \pietro{add ref}. Objects are simply maps
\begin{equation*}
    X \rightarrow F X,
\end{equation*}
where $X \in \Obj(\Cat)$. Morphisms are given by commutative diagrams
\begin{equation*}
    \begin{tikzcd}
        X \arrow[rightarrow]{d} \arrow[rightarrow]{r}{\phi}
        & Y \arrow[rightarrow]{d} \\
        FX \arrow[rightarrow]{r}{F \phi}
        & FY
    \end{tikzcd}
\end{equation*}
As $F$ is Cartesian, the category $\FCoalg$ is also Cartesian, with products given by
\begin{equation*}
    \prod_{i \in I} X_i \rightarrow \prod_{i \in I} F X_i \simeq F \prod_{i \in I} X_i.
\end{equation*}
As $F$ is nilpotent and preserves finite products, it is straightforward to verify that the forgetful functor
\begin{equation*}
    U \colon \FCoalg \rightarrow \Cat
\end{equation*}
has a right adjoint $\Phi$, called the cofree functor, given by
\begin{equation*}
    \Phi(X) = X \times F X \times \dots \times F^{n-1} X \rightarrow F X \times \dots \times F^{n-1} X.
\end{equation*}

\begin{remark}
    This adjunction is comonadic \pietro{explain! cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$.
\end{remark}

\section{Machines}

\begin{definition}\label{def:machine}
    Let $\Cat$ be a category, equipped with a terminal object $\pt$. Let $X \in \Obj(\Cat)$. A map
    \begin{equation*}
        \varrho \colon X \rightarrow X
    \end{equation*}
    is a {\em machine} if the induced transformation
    \begin{equation*}
        \Hom_\Cat(\pt, X) \rightarrow \Hom_\Cat(\pt, X)
    \end{equation*}
    has a unique fixed point $S_\varrho$, which we call the {\em stable state} of $\varrho$.
\end{definition}

\pietro{Fare qualche esempio, tipo contrazioni.}

\begin{definition}\label{def:parametric_machine}
    Let $\Cat$ be a category with finite products. A {\em parametric machine} is simply a machine on the Kleisli category for the comonad $P \times \anon$, for some $P \in \Obj(\Cat)$. More explicitly, a map
    \begin{equation*}
        \varrho \colon P \times X \rightarrow X
    \end{equation*}
    is a {\em machine} if there exists a unique map $S_\varrho$ such that the following diagram commutes
    \begin{equation}
        \begin{tikzcd}[column sep=large]\label[diagram]{diag:machine_condition}
            P \arrow[rightarrow]{r}{\left(\id, S_\varrho\right)}
            \arrow[rightarrow,swap]{dr}{S_\varrho}
            & P \times X \arrow[rightarrow]{d}{\varrho} \\
            & X
        \end{tikzcd}
    \end{equation}
\end{definition}

\Cref{diag:machine_condition} is analogous to the machine condition in~\cite{2020arXiv200702777V} in the case when $X$ is an Abelian group and $\varrho\colon P \times X \rightarrow X$ is the sum of a map $P \rightarrow X$ and a map $X \rightarrow X$. This approach differs from the one taken in~\cite{2020arXiv200702777V}, in that here the object $P$ represents both input space and parameters. In what follows, we will show how to construct examples of {\em machines} using {\em nilpotent functors}.

\begin{theorem}\label{thm:nilpotent_architecture}
    Let $F\colon \Cat \rightarrow \Cat$ be a nilpotent endofunctor. Let $\Phi X$ be a cofree coalgebra in $\FCoalg$. Let $P$ be a $F$ coalgebra. Then a map $P \times F X \rightarrow X$ induces a parametric machine $P \times \Phi X \rightarrow \Phi X$.
\end{theorem}

\begin{proof}
    \pietro{TODO per induzione.}
\end{proof}

\pietro{La sfida principale qui è trovare un buona architettura $F$ ma anche una buona coalgebra $P$. Probabilmente da menzionare da qualche parte che se la categoria di base è buona, $\FCoalg$ ha molte buone proprietà.}

\section{Recurrent and convolutional neural networks}

\pietro{discuss shape of filter?}

\section{Novel architectures}

\bibliographystyle{abbrv}
\bibliography{References}

\end{document}